Towards a miniKanren with fair search strategies

### 2019-05-25 Paper/paper.tex
rescue paper.tex

\NOTE[DPF]{Throughout the paper we have search spaces and search strategies.
I can see leaving in the  ``search strategies'', but let's try to
remove virtually all of the uses of ``search'' in front of spaces.}
\NOTE[LKC]{Corrected.}

\NOTE[LKC]{BUT, I am thinking about going further --- removing ``our improved'' 
from 
this paper. The first reason is that our BFS is actually more efficient than 
our improved BFS in most benchmarks (see Table 1), despite its theoretical 
space extravagence. The second reason is that DFSi, DFSf, and our BFS look nice 
when sitting together.}

\NOTE[DPF]{Does this mean that the comment about append goes away.
I really thought that was so clever.  I thought that the way you pointed
out the trivial changes was very clear, so that did not bother me at
all.}
\NOTE[LKC]{Yes, the comment about append will go away if we remove our improved 
BFS. Then, let's don't remove it, i.e. have two new BFS implementations at the 
same time.}
\NOTE[LKC]{I think having two new BFSs is a bit confusing. So I brand the 
new and easy/unimproved one as a ``stepping-stone'' and only mention it in BFS 
section.}

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can be
fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.  The two non-standard
depth-first searches are new. And we present a new, more efficient
and shorter implementation of the standard breadth-first search.
Using quantitative evaluation, we argue that a new depth-first searches is a
competitive alternative to the standard one, and that our improved 
breadth-first search implementation is more efficient than the
current one.

@@

\NOTE[DPF]{their BFS (or shortened to BFS). We need only say this once if we go 
with ``our'', which I don't like.
I like ``the new'' or ``the improved'' and ``the current'' }
\NOTE[LKC]{Corrected. I pick ``the improved''.}

@@

\NOTE[DPF]{The citetseres1999algebra must be not broken:
            You can either rewrite the sentence or
            use a tilde called a tie that means don't break the line.
            Sometimes using tilde does not work out elegant enough,
            so the first idea works best.}
\NOTE[LKC]{I have rewritten the whole item.}

@@

Now we elaborate fairness by running more queries about \repeato{}. We never 
use \runstar{} in this paper because fairness is more interesting when we 
have an unbounded number of answers. It is perfectly fine, however, to use 
\runstar{} with any search strategies.

@@

Under the hood, the \conde{} here is allocating computational resources to 
four trivially different spaces. The unfair \disj{} in 
\DFSi{} allocates many more resources to the first space. On the 
contrary, fair \disj{} would allocate resources evenly to each space.

@@

\DFSbi{} is fair only when the number of goals is a power of 2, 
otherwise, it allocates some goals twice as many resources as the 
others. In the above example, where the \conde{} has five clauses, \DFSbi{} 
allocates more resources to the clauses of \texttt{b}, \texttt{c}, and 
\texttt{d}.

@@

\NOTE[DPF]{Can you express exactly how all costs are determined?  I see that
for this simple program, your notion of cost makes sense, but how does it
generalize to all costs.}
\NOTE[LKC]{Corrected.}

@@

\NOTE[DPF]{I need clarification: Applying the above changes won't result in 
	``our \BFS,'' but yet a third implementation of \BFS?}
\NOTE[LKC]{If you still need clarification, please add a new note below. 
Otherwise, you might delete these two notes.}

\NOTE[DPF]{Why are the two texttt's macros, again?  Now, it hardly matters,
but it is unfriendly to your other co-authors to have us reading long
expressions when a newcommand would suffice.  I read the paper with a pen
marking it up, then I read the paper on the .tex screen.  That's each time
I read it.}
\NOTE[LKC]{I am sorry about that. Thank you for telling me! I have added many 
more macros and have removed many uses of texttt.}

\NOTE[DPF]{You know that ``bag'' is a technical terms.  It means that there
is a structure where each element of the structure has an associated number, 
which tells you have many occurrences of that element are in the list.
For example, the list (a b a c a d c) could be represented like this:
(a$^3$ b$^1$ c$^2$ d$^1$).  If that is not what you mean, you might change the 
name to bucket.}
\NOTE[LKC]{That is not what I mean. I have changed the name to bucket.}

@@

\NOTE[LKC]{I rewrite the second last sentence in the above paragraph such that 
it better corresponds to the comment about `append'.}

@@

\NOTE[DPF]{Font Space like in the code comments.}
\NOTE[LKC]{corrected}

We make the cost information more clear by changing the \texttt{Space} type, 
modifying related function definitions, and introducing a few more functions.

@@

\NOTE[DPF]{ks --> sk and kf --> fk and you don't need to say f and s
mean success continuation and failure continuation.}
\NOTE[LKC]{corrected}

Functions implementing impure features are in Fig.~\ref{BFSimp-cont}. The 
first function, \texttt{elim}, takes a space \sInf{} and two 
continuations \texttt{fk} and \texttt{sk}. When \sInf{} contains 
no answers, it calls \texttt{fk} with no argument. Otherwise, it calls 
\texttt{sk} with the first answer and the rest of the space. This function is 
similar to an eliminator of spaces, hence the name. The remaining 
functions are as before.

@@

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since \texttt{(run n ? g)} produces the same 
result in both \BFSimp{} and BFS. (See supplements for the formal proof.) To 
compare efficiency, we translate \BFSser{}'s Haskell code into Racket (See 
supplements for the translated code). The translation is direct due to the 
similarity of the two relational programming systems. The translated code is 
longer than \BFSimp{}. And it runs slower in all benchmarks. Details about 
differences in efficiency are in \autoref{sec:quan}.

@@

@@

\NOTE[DPF]{We don't have disequality constraints, so no absento.  You might just
say that.}
\NOTE[LKC]{Corrected.}

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} still performs more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}, the current standard miniKanren 
search strategy. 
Among the two implementations of \BFS, \BFSimp{} constantly perform as well 
or better. 
It is hard to conclude how \disj{} fairness affect performance based on these 
statistics.
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

% \BFSimp{}
We proof formally that \BFSimp{} is semantically equivalent to \BFSser{}. But, 
\BFSimp{} is shorter and performs better in comparison with a straightforward 
translation of their Haskell code.

>>>>

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can be
fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.  The two non-standard
depth-first searches are new. And we present a new, more efficient
and shorter implementation of the standard breadth-first search.
Using quantitative evaluation, we argue that each depth-first search is a
competitive alternative to the standard one, and that our improved 
breadth-first search implementation is more efficient than the
current one.

@@

\NOTE[DPF]{``current'' was for BFSser}

@@

@@

Now we elaborate fairness by running more queries about \repeato{}. We never 
use \runstar{} in this paper because fairness is more interesting when we 
have an unbounded number of answers. It is perfectly fine, however, to use 
\runstar{} with any search strategy.

@@

Under the hood, the \conde{} here allocates computational resources to 
four trivially different spaces. The unfair \disj{} in 
\DFSi{} allocates many more resources to the first space. On the 
contrary, fair \disj{} would allocate resources evenly to each space.

@@

\DFSbi{} is fair only when the number of goals is a power of 2, 
otherwise, it allocates some goals with twice as many resources as the 
others. In the above example, where the \conde{} has five clauses, \DFSbi{} 
allocates more resources to the clauses of \texttt{b}, \texttt{c}, and 
\texttt{d}.

@@

@@

@@

@@

We simplify the cost information by changing the \texttt{Space} type, 
modifying related function definitions, and introducing a few more functions.

@@

Functions implementing impure features are in Fig.~\ref{BFSimp-cont}. The 
first function, \texttt{elim}, takes a space \sInf{} and two 
continuations \texttt{fk} and \texttt{sk}. When \sInf{} contains 
no answers, it calls \texttt{fk}. Otherwise, it calls 
\texttt{sk} with the first answer and the rest of the space. This function is 
similar to an eliminator of spaces, hence the name. The remaining 
functions are as before.

@@

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since \texttt{(run n ? g)} produces the same 
result in both BFS and \BFSimp{}. (See supplements for the formal proof.) To 
compare efficiency, we translate \BFSser{}'s Haskell code into Racket (See 
supplements for the translated code). The translation is direct due to the 
similarity of the two relational programming systems. The translated code is 
longer than \BFSimp{}. And it runs slower in all benchmarks. Details about 
differences in efficiency are in \autoref{sec:quan}.

\NOTE[DPF]{Make another macro or use a number when we want to drop ``the''
from ``the improved BFS''.  In the table below, use that macro.}

@@

\NOTE[DPF]{tilde problem 7 lines down.}

@@

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}, the current standard miniKanren 
search strategy. 
Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
or better. 
It is hard to conclude how \disj{} fairness affects performance based on these 
statistics.
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

% \BFSimp{}
We prove formally that \BFSimp{} is semantically equivalent to \BFSser{}. But, 
\BFSimp{} is shorter and performs better in comparison with a straightforward 
translation of their Haskell code.

\NOTE[DPF]{Please leave ``prove'' and don't use ``proof''.}
